Cycles of the logistic map

### 2013-06-03 doc/pc.tex
add lsfit.ma, T14 polynomials

Consider the logistic map \cite{may, strogatz}:
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv r \, x_k \, ( 1 - x_k ).
\label{eq:logmap}
\end{equation}
%
The iterated \refeq{logmap} sequence $x_1$,
  $x_2 = f(x_1)$,
  $x_3 = f(x_2)$, $\ldots$
can be conveniently visualized on the cobweb plot,
  as shown in \reffig{cobweb}.
%
Starting from $(x_1, x_1)$ on the diagonal,
  each vertical arrow takes $(x_k, x_k)$ to $(x_k, y)$,
  where $y = x_{k+1} = f(x_k)$;
the next horizontal arrow then reflects $(x_k, y)$ to
  $(y, y) = (x_{k+1}, x_{k+1})$,
  the starting point of the next iteration.
%
\begin{figure}[h]
  \begin{minipage}{\linewidth}
    \begin{center}
        \includegraphics[angle=-90, width=\linewidth]{cobweb.pdf}
    \end{center}
  \end{minipage}%
  \caption{\label{fig:cobweb}
  Cobweb plot of the logistic map.}
\end{figure}
%
%
After many iterations, there can be three ultimate patterns:
  (i) a \emph{fixed point} or a constant (including infinity),
      e.g., \reffig{cobweb}(a);
  (ii) a periodic \emph{cycle} or a self-repeating loop,
      e.g., \reffigs{cobweb}(b)-(e);
or
  (iii) a chaotic trajectory,
      e.g., \reffig{cobweb}(f).
We will focus on the first two cases here.

A fixed point is a solution of $x^* = f(x^*)$.
%
It is stable if $x_1$ deviates slightly from $x^*$
  and the sequence still converges to $x^*$.
%
For a differentiable $f$,
  a stable fixed point requires $|f'(x^*)| \le 1$
  to reduce deviations in successive iterations \cite{strogatz}.

An $n$-cycle is a sequence $x_1, \dots, x_n$,
  such that $n$ is the smallest positive integer
  that satisfies $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
Obviously, any $x_k$ in an $n$-cycle of $f$ is a fixed point of $f^n$
%
\big[but the converse is untrue, for a fixed point of $f^n$
  can also be a fixed point of $f^d$ for a divisor $d$ of $n$:
  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big].
%
%
Thus, a cycle can be classified as stable or unstable
  according to the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_1) \cdots f'(x_n) \Big| \le 1,
\label{eq:der}
\end{equation}
%
%
%The set of the $n$ points within the cycle $x_1. \ldots, x_n$
%   is called the \emph{orbit}.
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz}.
%
%
%

@@

\subsection{\label{sec:primfac}Minimal polynomial for the $n$-cycles}

@@

Each fixed point can be assigned to
  a point in a $d$-cycle, with $d$ being a divisor of $n$.
%
The assignment
  is both complete (for a $d$-cycle point
  must also be a fixed point of $f^n$)
  and non-redundant (for there is no
  repeated fixed point of $f^n$).
%
Since each of the $L(d)$ $d$-cycles
  contributes $d$ fixed points,
we have
$2^n = \sum_{d\,|\,n} L(d) \, d$.
%
The M\"obius inversion yields $L(n) = (1/n) \sum_{d\,|\,n} \mu(n/d) 2^{\,d}$.
This formula was known to several authors \cite{hao, lutzky}.

@@

@@

\bibitem{hao} B.-L. Hao,
  \textit{Elementary Symbolic Dynamics and Chaos in  Dissipative Systems}
  (World Scientific, Singapore, 1989);
  Number of periodic orbits in continuous maps of
    the interval complete solution of the counting problem,
  \textit{Annals of Combinatorics}
  \textbf{4},
  339--346
  (2000).

>>>>

Consider the logistic map \cite{may, strogatz}:
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv r \, x_k \, ( 1 - x_k ).
\label{eq:logmap}
\end{equation}
%
The iterated \refeq{logmap} sequence $x_1$,
  $x_2 = f(x_1)$,
  $x_3 = f(x_2)$, $\ldots$
can be conveniently visualized on the cobweb plot,
  as shown in \reffig{cobweb}.
%
Starting from $(x_1, x_1)$ on the diagonal,
  each vertical arrow takes $(x_k, x_k)$ to $(x_k, y)$,
  where $y = x_{k+1} = f(x_k)$;
the next horizontal arrow then reflects $(x_k, y)$ to
  $(y, y) = (x_{k+1}, x_{k+1})$,
  the starting point of the next iteration.
%
\begin{figure}[h]
  \begin{minipage}{\linewidth}
    \begin{center}
        \includegraphics[angle=-90, width=\linewidth]{cobweb.pdf}
    \end{center}
  \end{minipage}%
  \caption{\label{fig:cobweb}
  Cobweb plot of the logistic map.}
\end{figure}
%
%
After many iterations, there can be three ultimate patterns:
  (i) a \emph{fixed point} or a constant (including infinity),
      e.g., \reffig{cobweb}(a);
  (ii) a periodic \emph{cycle} or a self-repeating loop,
      e.g., \reffigs{cobweb}(b)-(e);
or
  (iii) a chaotic or aperiodic trajectory,
      e.g., \reffig{cobweb}(f).
We will focus on the first two cases here.

A fixed point is a solution of $x^* = f(x^*)$.
%
It is stable if $x_1$ deviates slightly from $x^*$
  and the sequence still converges to $x^*$.
%
For a differentiable $f$,
  a stable fixed point requires $|f'(x^*)| \le 1$
  to reduce deviations in successive iterations \cite{strogatz, hao}.

An $n$-cycle is a sequence $x_1, \dots, x_n$,
  such that $n$ is the smallest positive integer
  that satisfies $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
Obviously, any $x_k$ in an $n$-cycle of $f$ is a fixed point of $f^n$
%
\big[but the converse is untrue, for a fixed point of $f^n$
  can also be a fixed point of $f^d$ for a divisor $d$ of $n$:
  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big].
%
%
Thus, a cycle can be classified as stable or unstable
  according to the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_1) \cdots f'(x_n) \Big| \le 1,
\label{eq:der}
\end{equation}
%
%
%The set of the $n$ points within the cycle $x_1. \ldots, x_n$
%   is called the \emph{orbit}.
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz, hao}.
%
%
%

@@

\subsection{\label{sec:primfac}Polynomials of $n$-cycles}

@@

Each fixed point can be assigned to
  a point in a $d$-cycle, with $d$ being a divisor of $n$.
%
The assignment
  is both complete (for a $d$-cycle point
  must also be a fixed point of $f^n$)
  and non-redundant (for there is no
  repeated fixed point of $f^n$).
%
Since each of the $L(d)$ $d$-cycles
  contributes $d$ fixed points,
we have
$2^n = \sum_{d\,|\,n} L(d) \, d$.
%
The M\"obius inversion yields $L(n) = (1/n) \sum_{d\,|\,n} \mu(n/d) 2^{\,d}$.
This formula was known to several authors \cite{hao2, lutzky}.

@@

\bibitem{hao} B.-L. Hao,
  \textit{Elementary Symbolic Dynamics and Chaos in  Dissipative Systems}
  (World Scientific, Singapore, 1989).

@@

\bibitem{hao2} B.-L. Hao,
  Number of periodic orbits in continuous maps of
    the interval complete solution of the counting problem,
  \textit{Annals of Combinatorics}
  \textbf{4},
  339--346
  (2000).
